---
title: "R Modeling Notebook for Practicum Project, 2023"
output: html_notebook
---
Social Determinants of Health, Air Quality, and the Prediction of Respiratory Disease

```{r}
#Load necessarily libraries, set seed, read in data, etc. 

#Load libraries
library(MASS)
library(xgboost)
library(writexl)
library(fitdistrplus)
library(nortest)
library(ggplot2)
library(broom)
library(corrplot)
library(randomForest)
library(olsrr)
library(mixtools)
library(glmnet)
library(stats)
library(caret)
library("FactoMineR")
library(factoextra)
library(dplyr)
library(rpart) #for fitting decision trees
library(rpart.plot)
library(ClusterR)

#Set Seed
SEED = 10
set.seed(SEED)

#Read in data
DF_Total <-read.csv("Final_DF_For_Modeling.csv", header=TRUE)
DF_Total
#Drop contextual variables like fips codes or names and varaibles determined to be highly correlated by VIF and heatmap
#
DF <- subset(DF_Total, select = -c(CountyFips, County_Name_Short, State_Accnym, State_Long, County_Long, E_HH, E_HU, Number.of.business.establishments,Air.Quality.Bill.Count..current..recent..and.prior..80.percent.match.and.above.for..air.quality., Percent.Properties.Low.Fire.Risk, Percent_Under_5,Percent_5_to_14,Percent_15_to_24,Percent_25_to_34,Percent_35_to_44, Percent_45_to_54, Percent_55_to_64, Percent_65_to_74, Percent_75_and_Up, Male.2022...of.Pop.))

#Prevalence.Pediatric.Asthma, Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer

#Convert type (some were char) after stripping out the percent sign 
#aslo had to manually remove a couple "#DIV/0!" in excel because they should have been zeros

#Sum Medium or high fire risk properties and drop individual columns
DF$Percent.Properties.Medium.Or.High.Fire.Risk <- DF$Percent.Properties.High.Fire.Risk+DF$Percent.Properties.Medium.Fire.Risk

DF <- subset(DF, select = -c(Percent.Properties.Medium.Fire.Risk, Percent.Properties.High.Fire.Risk))

DF[] <- lapply(DF, function(x) as.numeric(gsub("%", "", x)))

#checking for NA, there are none
#summary(is.na(DF))


#Scale data, put col names back, and then shift so between 0 and 1 (this is needed for PCA)

DF_Scaled_1 <- scale(DF, center=TRUE, scale=TRUE)
DF_Scaled1 <- as.data.frame(DF_Scaled_1)
colnames(DF_Scaled1) <- colnames(DF)
DF_Scaled <- (DF_Scaled1 - min(DF_Scaled1)) / (max(DF_Scaled1) - min(DF_Scaled1))

DF_Scaled_Without_Predictors <- subset(DF_Scaled, select = -c(Prevalence.Pediatric.Asthma, Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))

scaleList <- list(scale = attr(DF_Scaled_1, "scaled:scale"),
    center = attr(DF_Scaled_1, "scaled:center"))

```


```{r}

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.


Helpful References:
https://www.statology.org/test-for-normality-in-r/
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3693611/
https://www.statology.org/shapiro-wilk-test-r/
https://towardsdatascience.com/your-data-isnt-normal-54fe98b1f322
https://stats.stackexchange.com/questions/232011/ties-should-not-be-present-in-one-sample-kolmgorov-smirnov-test-in-r

https://cran.r-project.org/web/packages/olsrr/vignettes/intro.html
https://www.statology.org/ridge-regression-in-r/

Kolmogorov-Smirnov Test:
If the p-value is below 0.05 then it indicates the data isn't normally distributed. If it is above 0.05 then it indicates a normal distribution.
Unfortunatley not a great way to test for normalacy

```{r}
DF_No_Predictors <- subset(DF, select = -c(Prevalence.Pediatric.Asthma, Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))

Predictors <- subset(DF, select = c(Prevalence.Pediatric.Asthma, Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))

DF_Linear_Regression <- scale(DF_No_Predictors, center=TRUE, scale=TRUE)
colnames(DF_Linear_Regression) <- colnames(DF_No_Predictors)

DF_Lin <- cbind(DF_Linear_Regression, Predictors)

set.seed(123)
#training test
n <- nrow(DF)
sample_indices <- sample(1:n, size = n, replace = FALSE)
train_size <- round(0.7 * n)
test_size <- n - train_size
train_data_Lin <- DF_Lin[sample_indices[1:train_size], ]
test_data_Lin <- DF_Lin[sample_indices[(train_size + 1):n], ]

train <-subset(train_data_Lin, select = -c(Prevalence.Pediatric.Asthma, Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
train_response <- subset(train_data_Lin, select = c(Prevalence.Pediatric.Asthma, Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
test <- subset(test_data_Lin, select = -c(Prevalence.Pediatric.Asthma, Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
test_response <- subset(test_data_Lin, select = c(Prevalence.Pediatric.Asthma, Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))

train_data_CL_COPD <-subset(train_data_Lin, select = -c(Prevalence.Pediatric.Asthma, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
train_data_CL_PA <-subset(train_data_Lin, select = -c(Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
train_data_CL_AA <-subset(train_data_Lin, select = -c(Prevalence.COPD, Prevalence.Pediatric.Asthma, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer))
train_data_CL_LC <-subset(train_data_Lin, select = -c(Prevalence.COPD, Prevalence.Pediatric.Asthma, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Adult.Asthma))
train_data_CL_CLD <-subset(train_data_Lin, select = -c(Prevalence.COPD, Prevalence.Pediatric.Asthma, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))

#check for normal distribution of residuals of all possible response variables


all_data_COPD <-subset(DF_Lin, select = -c(Prevalence.Pediatric.Asthma, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
all_data_PA <-subset(DF_Lin, select = -c(Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
all_data_AA <-subset(DF_Lin, select = -c(Prevalence.COPD, Prevalence.Pediatric.Asthma, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer))
all_data_LC <-subset(DF_Lin, select = -c(Prevalence.COPD, Prevalence.Pediatric.Asthma, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Adult.Asthma))
all_data_CLD <-subset(DF_Lin, select = -c(Prevalence.COPD, Prevalence.Pediatric.Asthma, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))


```


```{r}

#og training and test r-squared and rmse

PA_OG <- lm(train_data_CL_PA$Prevalence.Pediatric.Asthma~., data = train_data_CL_PA)
summary(PA_OG)
predictions <- predict(PA_OG, newdata = test)
RMSE(predictions,test_response$Prevalence.Pediatric.Asthma)

AA_OG <-lm(train_data_CL_AA$Prevalence.Adult.Asthma~., data = train_data_CL_AA)
summary(AA_OG)
predictions <- predict(AA_OG, newdata = test)
RMSE(predictions,test_response$Prevalence.Adult.Asthma)

COPD_OG <- lm(train_data_CL_COPD$Prevalence.COPD~., data = train_data_CL_COPD)
summary(COPD_OG)
predictions <- predict(COPD_OG, newdata = test)
RMSE(predictions,test_response$Prevalence.COPD)

CLD_OG <- lm(train_data_CL_CLD$Prevalence.Adult.Chronic.Lung.Disease~., data = train_data_CL_CLD)
summary(CLD_OG)
predictions <- predict(CLD_OG, newdata = test)
RMSE(predictions,test_response$Prevalence.Adult.Chronic.Lung.Disease)

COPD_OG_transformed <- lm(log(train_data_CL_COPD$Prevalence.COPD)^2~., data = train_data_CL_COPD)
summary(COPD_OG_transformed)
predictions <- predict(COPD_OG_transformed, newdata = test)
val <- exp(sqrt(predictions))
RMSE(val,test_response$Prevalence.COPD)

```





```{r}
#Assumption check
PA <- lm(all_data_PA$Prevalence.Pediatric.Asthma~., data = all_data_PA)
AA <-lm(all_data_AA$Prevalence.Adult.Asthma~., data = all_data_AA)
COPD <- lm(all_data_COPD$Prevalence.COPD~., data = all_data_COPD)
CLD <- lm(all_data_CLD$Prevalence.Adult.Chronic.Lung.Disease~., data = all_data_CLD)
LC <- lm(all_data_LC$Prevalence.Lung.Cancer~., data = all_data_LC)

### Histograms

hist(PA$residuals, col= "darkblue", nclass = 15, main = "Pediatric Asthma Residual Distribution") 
hist(AA$residuals, col="yellow", main = "Adult Asthma Residual Distribution") 
hist(COPD$residuals, col= "darkgreen", main = "COPD Residual Distribution") 
hist(CLD$residuals, col= "darkorange", main = "Adult Chronic Lung Disease Residual Distribution") 
hist(LC$residuals, col= "lightblue", main = "Lung Cancer Residual Distribution") 


####QQ Plots
qqnorm(PA$residuals, col= "darkblue", main='Pediatric Asthma Residual Distribution')
qqline(PA$residuals)
qqnorm(AA$residuals, col="Gold",main='Adult Asthma Residual Distribution')
qqline(AA$residuals)
qqnorm(COPD$residuals, col= "darkgreen",main='COPD Residual Distribution')
qqline(COPD$residuals)
qqnorm(CLD$residuals, col= "darkorange",main='Adult Chronic Lung Disease Residual Distribution')
qqline(CLD$residuals)
qqnorm(LC$residuals, col= "lightblue",main='Lung Cancer Residual Distribution')
qqline(LC$residuals)


#ggplot(data=DF_Scaled, aes(x=Lin_reg_assump$fitted.values, y=Lin_reg_assump$residuals)) + #geom_point(alpha=I(0.4),color='darkorange') +
#xlab('Fitted Values') +
#ylab('Residuals') +
#ggtitle('COPD Residual Plot') + geom_hline(yintercept=0)


#normal_data <- rnorm(200)


##Cullen and Frey Graph

descdist(PA$residuals, discrete = FALSE, boot = NULL, method = "unbiased", graph = TRUE, obs.col = "darkblue", obs.pch = 16, boot.col = "orange")
descdist(AA$residuals, discrete = FALSE, boot = NULL, method = "unbiased", graph = TRUE, obs.col = "yellow", obs.pch = 16, boot.col = "orange")
descdist(COPD$residuals, discrete = FALSE, boot = NULL, method = "unbiased", graph = TRUE, obs.col = "darkgreen", obs.pch = 16, boot.col = "orange")
descdist(CLD$residuals, discrete = FALSE, boot = NULL, method = "unbiased", graph = TRUE, obs.col = "darkorange", obs.pch = 16, boot.col = "orange")
descdist(LC$residuals, discrete = FALSE, boot = NULL, method = "unbiased", graph = TRUE, obs.col = "lightblue", obs.pch = 16, boot.col = "orange")


#uncorrelated errors and constant variance checks
ggplot(data=all_data_PA, aes(x=PA$fitted.values, y=PA$residuals)) + geom_point(alpha=I(0.4),color='darkblue') +
xlab('Fitted Values') + ylab('Residuals') + ggtitle('Pediatric Asthma Residual Plot') + geom_hline(yintercept=0)


###another way to do cooks distance
#plot(AA, 4)
###For future reference a good way to get diagnostic plots quickly
#ols_plot_diagnostics(COPD)
```


```{r}
###Cook's Distance

plot(PA, 5)


cooksd <- cooks.distance(PA)
# Plot
plot(cooksd, pch = 19, type = "h", main = "Cook's Distance PA")

# Higher than 0.5 is worth looking at, higher than 1 is very high
high_cooksd_rows <- which(cooksd > .2)

# Trace back to original county and state to look up manually (Loving and Kenedy chosen as examples)
print(DF_Total[high_cooksd_rows, ])

###PA, AA, COPD, CLD, LC
```


```{r}
#Get a pdf image of a correlation heat map for variables 
Get_Correlations <- cor(DF_Scaled)
pdf("correlation_plot.pdf", width = 10, height = 10)  # Adjust width and height as needed

# Create the correlation plot (heatmap)
corrplot(Get_Correlations, method = "color", type = "lower", order = "hclust", tl.cex = 0.6, tl.col = "black")
dev.off()

#multicolinearity check
multicollin <- ols_coll_diag(AA)
multicollin$vif_t

multicollin$vif_t

write.csv(multicollin$vif_t, file = "vif.csv", row.names = FALSE)

```


```{r}
### Model Transformations
### did not help except for COPD

##Lung Cancer attempt
log_y <- (DF_Lin$Prevalence.Lung.Cancer)^2

# Code to calculate lambda
# Extract optimal lambda
#^all done on fully scaled data to avoid negative values

LC2 <- lm(log_y~., data = all_data_LC)
qqnorm(LC2$residuals, main='Lung Cancer Transformed')
qqline(LC2$residuals)


### Transformation for COPD and PA

log_y <- log(DF_Lin$Prevalence.Pediatric.Asthma)^2

norespon <- subset(DF_Lin, select = -c(Prevalence.Pediatric.Asthma, Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))

  
PA2 <- lm(log_y~., data = norespon)
qqnorm(PA2$residuals, main='COPD Transformed')
qqline(PA2$residuals)


descdist(PA2$residuals, discrete = FALSE, boot = NULL, method = "unbiased", graph = TRUE, obs.col = "darkblue", obs.pch = 16, boot.col = "orange")
############

log_y <- log(DF_Lin$Prevalence.COPD)^2

norespon <- subset(DF_Lin, select = -c(Prevalence.Pediatric.Asthma, Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))

  
COPD2 <- lm(log_y~., data = norespon)
qqnorm(COPD2$residuals, main='COPD Transformed')
qqline(COPD2$residuals)

descdist(COPD2$residuals, discrete = FALSE, boot = NULL, method = "unbiased", graph = TRUE, obs.col = "darkgreen", obs.pch = 16, boot.col = "orange")

```


"Note that setting alpha equal to 1 is equivalent to using Lasso Regression and setting alpha to some value between 0 and 1 is equivalent to using an elastic net."
https://www.statology.org/ridge-regression-in-r/


```{r}


####Variable selection

#Originally 34 variables

#backward/forward
backward_vs <- step(COPD, direction = "backward", trace = 0)
selected_variables <- attr(terms(backward_vs), "term.labels")
new_data_selected <- DF_Scaled[, selected_variables]
# Use selected variables in a new modeling attempt to see R squared improvement
new_model <- lm((log(DF_Scaled$Prevalence.COPD))^2 ~ ., data = new_data_selected)
summary(new_model)




# Extract all variable names from your data frame
all_variables <- colnames(DF_Scaled_Without_Predictors)

# Identify variables not selected
not_selected_variables <- setdiff(all_variables, selected_variables)

# List the variables that were not selected
cat("Number of variables selected:", length(selected_variables), "\n")
cat("Variables not selected:", not_selected_variables, "\n")

#########Lasso/Ridge/Elastic

set.seed(10)
#when alpha 1, lasso, check if reduces number of var
# Fit a Ridge regression model with cross-validation
X_matrix <- as.matrix(norespon)
#Y_matrix <- as.matrix(((log(DF_Scaled$Prevalence.COPD))^2))
Y_matrix <- as.matrix(DF_Lin$Prevalence.Adult.Asthma)
#fit ridge regression model
model <- glmnet(X_matrix, Y_matrix, alpha = 0)
#view summary of model
summary(model)
#perform k-fold cross-validation to find optimal lambda value
cv_model <- cv.glmnet(X_matrix, Y_matrix, alpha = 0)
#find optimal lambda value that minimizes test MSE
best_lambda <- cv_model$lambda.min
#produce plot of test MSE by lambda value
plot(cv_model) 
#find coefficients of best model
best_model <- glmnet(X_matrix, Y_matrix, alpha = 0, lambda = best_lambda)
coef(best_model)
y_predicted <- predict(model, s = best_lambda, newx = X_matrix)
#find SST and SSE
sst <- sum((Y_matrix - mean(Y_matrix))^2)
sse <- sum((y_predicted - Y_matrix)^2)
#find R-Squared
rsq <- 1 - sse/sst
rsq




```

SHAP
https://cran.r-project.org/web/packages/shapr/vignettes/understanding_shapr.html


https://www.datacamp.com/tutorial/pca-analysis-r

```{r}
###PCA

data.pca <- princomp(DF_Scaled_Without_Predictors)
summary(data.pca)
data.pca$loadings[, 1:2]
data.pca$loadings

fviz_eig(data.pca, addlabels = TRUE)


fviz_pca_var(data.pca, col.var = "black", )

fviz_pca_var(data.pca, col.var = "cos2",
            gradient.cols = c("black", "orange", "green"),
            repel = TRUE)

fviz_cos2(data.pca, choice = "var", axes = 1:3)




set.seed(10)

# Perform PCA on the predictors
predictors <- as.matrix(DF_Scaled_Without_Predictors)
pca_result <- prcomp(predictors, center = TRUE, scale = TRUE)

Y <- as.matrix(DF_Scaled$Prevalence.Adult.Chronic.Lung.Disease)

# Extract the principal components
principal_components <- pca_result$x
# Combine the principal components and the response variable into a new data frame
pca_data_frame <- data.frame(
  PC1 = principal_components[, 1],
  PC2 = principal_components[, 2],
  PC3 = principal_components[, 3],
  PC4 = principal_components[, 4],
  PC5 = principal_components[, 5],
  PC6 = principal_components[, 6],
  PC7 = principal_components[, 7],
  PC8 = principal_components[, 8],
  PC9 = principal_components[, 9],
  PC10 = principal_components[, 10],
  PC11 = principal_components[, 11],
  PC12 = principal_components[, 12],
  PC13 = principal_components[, 13],
  PC14 = principal_components[, 14],
  PC15 = principal_components[, 15],
  PC16 = principal_components[, 16],
  PC17 = principal_components[, 17],
  PC18 = principal_components[, 18],
  PC19 = principal_components[, 19],
  PC20 = principal_components[, 20],
  PC21 = principal_components[, 21],
  PC22 = principal_components[, 22],
  PC23 = principal_components[, 23],
  PC24 = principal_components[, 24],
  PC25 = principal_components[, 25],
  PC26 = principal_components[, 26],
  PC27 = principal_components[, 27],
  PC28 = principal_components[, 28],
  PC29 = principal_components[, 29],
  PC30 = principal_components[, 30],
  PC31 = principal_components[, 31],
  PC32 = principal_components[, 32],
  PC33 = principal_components[, 33]
)

# Fit a linear regression model using the principal components
model <- lm(Y ~ ., data = pca_data_frame)
summary(model)
```


```{r}
#### Training and testing#########

train <-subset(train_data_Lin, select = -c(Prevalence.Pediatric.Asthma, Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
train_response <- subset(train_data_Lin, select = c(Prevalence.Pediatric.Asthma, Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
test <- subset(test_data_Lin, select = -c(Prevalence.Pediatric.Asthma, Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
test_response <- subset(test_data_Lin, select = c(Prevalence.Pediatric.Asthma, Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))

train_data_CL_COPD <-subset(train_data_Lin, select = -c(Prevalence.Pediatric.Asthma, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
train_data_CL_PA <-subset(train_data_Lin, select = -c(Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
train_data_CL_AA <-subset(train_data_Lin, select = -c(Prevalence.COPD, Prevalence.Pediatric.Asthma, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer))
train_data_CL_LC <-subset(train_data_Lin, select = -c(Prevalence.COPD, Prevalence.Pediatric.Asthma, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Adult.Asthma))
train_data_CL_CLD <-subset(train_data_Lin, select = -c(Prevalence.COPD, Prevalence.Pediatric.Asthma, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))


######Get the step models 

########COPD
backward_copd <- step(COPD2, direction = "backward", trace = 0)
selected_variables <- attr(terms(backward_vs), "term.labels")
new_data_selected <- train_data_CL_COPD[, selected_variables]
# Use selected variables in a new modeling attempt to see R squared improvement
backward_copd_model_train <- lm((log(train_data_CL_COPD$Prevalence.COPD))^2 ~ ., data = new_data_selected)
summary(backward_copd_model_train)

predictionsCOPD <- predict(backward_copd_model_train, newdata = test)
predictions_original_scale <- exp(sqrt(predictionsCOPD))

RMSE(predictions_original_scale,test_response$Prevalence.COPD)

#OG COPD

backward_copd_OG <- step(COPD, direction = "backward", trace = 0)
selected_variables <- attr(terms(backward_copd_OG), "term.labels")
new_data_selected <- train_data_CL_COPD[, selected_variables]

# Use selected variables in a new modeling attempt to see R squared improvement
backward_copd_model_train_OG <- lm(train_data_CL_COPD$Prevalence.COPD ~ ., data = new_data_selected)
summary(backward_copd_model_train_OG)

predictionsCOPD <- predict(backward_copd_model_train_OG, newdata = test)

RMSE(predictionsCOPD,test_response$Prevalence.COPD)

mean(test_response$Prevalence.COPD)


########CLD

backward_cld <- step(CLD, direction = "backward", trace = 0)
selected_variables <- attr(terms(backward_vs), "term.labels")
new_data_selected <- train_data_CL_CLD[, selected_variables]
# Use selected variables in a new modeling attempt to see R squared improvement
backward_cld_model_train <- lm(train_data_CL_CLD$Prevalence.Adult.Chronic.Lung.Disease ~ ., data = new_data_selected)
summary(backward_cld_model_train)

predictions_cld <- predict(backward_cld_model_train, newdata = test)

RMSE(predictions_cld,test_response$Prevalence.Adult.Chronic.Lung.Disease)
mean(test_response$Prevalence.Adult.Chronic.Lung.Disease)

########AA

backward_aa <- step(AA, direction = "backward", trace = 0)
selected_variables <- attr(terms(backward_vs), "term.labels")
new_data_selected <- train_data_CL_AA[, selected_variables]
# Use selected variables in a new modeling attempt to see R squared improvement
backward_aa_model_train <- lm(train_data_CL_AA$Prevalence.Adult.Asthma ~ ., data = new_data_selected)
summary(backward_aa_model_train)

predictions_aa <- predict(backward_aa_model_train, newdata = test)

RMSE(predictions_aa,test_response$Prevalence.Adult.Asthma)


mean(test_response$Prevalence.Adult.Asthma)


########PA
backward_pa <- step(PA, direction = "backward", trace = 0)
selected_variables <- attr(terms(backward_vs), "term.labels")
new_data_selected <- train_data_CL_PA[, selected_variables]
# Use selected variables in a new modeling attempt to see R squared improvement
backward_pa_model_train <- lm(train_data_CL_PA$Prevalence.Pediatric.Asthma ~ ., data = new_data_selected)
summary(backward_pa_model_train)

predictions_pa <- predict(backward_pa_model_train, newdata = test)

RMSE(predictions_pa,test_response$Prevalence.Pediatric.Asthma)

mean(test_response$Prevalence.Pediatric.Asthma)

mean(test_response$Prevalence.Lung.Cancer)

#best model, get equation and examples

model_tidy <- tidy(backward_copd_model_train_OG)
model_tidy$estimate


predictionsCOPD <- predict(backward_copd_model_train_OG, newdata = test_data_Without_Response_copd)

RMSE(predictionsCOPD,test_data_Response_copd$Prevalence.COPD)
```


```{r}
#Redo to see without scale
COPD_NO_0_1_Scale <- lm(DF_Scaled1$Prevalence.COPD~., data = DF_Scaled1)
set.seed(123)
#training test
n <- nrow(DF_Scaled)
sample_indices <- sample(1:n, size = n, replace = FALSE)
train_size <- round(0.7 * n)
test_size <- n - train_size
train_data2 <- DF_Scaled1[sample_indices[1:train_size], ]
train_data2 <- subset(train_data2, select = -c(Prevalence.Pediatric.Asthma, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
test_data2 <- DF_Scaled1[sample_indices[(train_size + 1):n], ]
test_data2 <- subset(test_data2, select = -c(Prevalence.Pediatric.Asthma, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
#backwards
backward_copd_OG_2 <- step(COPD_NO_0_1_Scale, direction = "backward", trace = 0)
selected_variables <- attr(terms(backward_copd_OG_2), "term.labels")
new_data_selected <- DF_Scaled1[, selected_variables]

# Use selected variables in a new modeling attempt to see R squared improvement
BW_COPD_Scale <- lm(train_data2$Prevalence.COPD ~ ., data = train_data2)
summary(BW_COPD_Scale)

singlepoint <- data.frame(test_data2[7, ])
singlepoint
p <- predict(BW_COPD_Scale, newdata = singlepoint)
p

p_converted_back_to_OG_scale <- p * scaleList$scale["Prevalence.COPD"] + scaleList$center["Prevalence.COPD"]
p_converted_back_to_OG_scale


Singlepoint_converted_back <- singlepoint$Prevalence.COPD * scaleList$scale["Prevalence.COPD"] + scaleList$center["Prevalence.COPD"]
Singlepoint_converted_back

#redo coefficients
model_tidy <- tidy(BW_COPD_Scale)
model_tidy$estimate
model_tidy$term

predictionsCOPD <- predict(backward_copd_model_train_OG, newdata = test_data_Without_Response_copd)

RMSE(predictionsCOPD,test_data_Response_copd$Prevalence.COPD)
```



```{r}
#Get together Non-linear modeling data (No scale)

set.seed(123)
#training test
n <- nrow(DF)
sample_indices <- sample(1:n, size = n, replace = FALSE)
train_size <- round(0.7 * n)
test_size <- n - train_size
train_data_NL <- DF[sample_indices[1:train_size], ]
test_data_NL <- DF[sample_indices[(train_size + 1):n], ]

train_data_NL

#get response
COPD_Train <-subset(train_data_NL, select = -c(Prevalence.Pediatric.Asthma, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
COPD_Test <- subset(test_data_NL, select = -c(Prevalence.Pediatric.Asthma, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
PA_Train <- subset(train_data_NL, select = -c(Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma, Prevalence.COPD))
PA_Test <- subset(test_data_NL, select = -c(Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma, Prevalence.COPD))
AA_Train <- subset(train_data_NL, select = -c(Prevalence.Pediatric.Asthma, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.COPD))
AA_Test <- subset(test_data_NL, select = -c(Prevalence.Pediatric.Asthma, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.COPD))
CLD_Train <- subset(train_data_NL, select = -c(Prevalence.Pediatric.Asthma, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma, Prevalence.COPD))
CLD_Test <- subset(test_data_NL, select = -c(Prevalence.Pediatric.Asthma, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma, Prevalence.COPD))
LC_Train <- subset(train_data_NL, select = -c(Prevalence.Pediatric.Asthma, Prevalence.Adult.Chronic.Lung.Disease,Prevalence.Adult.Asthma, Prevalence.COPD))
LC_Test <- subset(test_data_NL, select = -c(Prevalence.Pediatric.Asthma, Prevalence.Adult.Chronic.Lung.Disease,Prevalence.Adult.Asthma, Prevalence.COPD))

#remove response
#train_data_NL <- subset(train_data_NL, select = -c(Prevalence.Pediatric.Asthma, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma, Prevalence.COPD))

#test_data_NL <- subset(test_data_NL, select = -c(Prevalence.Pediatric.Asthma, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma, Prevalence.COPD))
```


```{r}
#### Random Forest

# Fit a Random Forest model
set.seed(123)  # For reproducibility
rf_model <- randomForest(CLD_Train$Prevalence.Adult.Chronic.Lung.Disease ~ ., data = CLD_Train, ntree = 100, importance = TRUE)

# Print a summary of the Random Forest model
print(rf_model)

# Assess variable importance
var_importance <- round(importance(rf_model), 2)
sorted_var_importance <- var_importance[order(-var_importance[, 1]), ]

# Print sorted variable importance
cat("Variable Importance (sorted):\n")
print(sorted_var_importance)

# Calculate the R-squared value

# Make predictions (change "new_data" to your new data for prediction)
# new_data <- data.frame(...)  # Replace with your new data
predictions <- predict(rf_model, newdata = CLD_Test)

Test_observations <- CLD_Test$Prevalence.Adult.Chronic.Lung.Disease
# Evaluate the model's performance (e.g., using confusion matrix, ROC, etc.)
# Calculate the R-squared value

rsquared <- R2(pred = predictions, obs = Test_observations)
print(paste("R-squared:", rsquared))

# RMSE
RMSE(predictions,Test_observations)
```

https://www.statology.org/classification-and-regression-trees-in-r/

```{r}
#Reg Trees
#LC

#CHANGE THESE VARIABLES ONCE DONE UPDATING
RT_Limited_Variables_Train <- subset(train_data_NL, select = c(Number.Policy.Passed, Percent.Policy.Passed.Of.All.Proposed, Number.of.Allergen.Species, Smoking.prevalence..as.a.percentage.of.the.population., E_TOTPOP, High.Pollen.State., Percent.Properties.Medium.Or.High.Fire.Risk, EP_UNINSUR, Prevalence.Lung.Cancer))

RT_Limited_Variables_Test <- subset(train_data_NL, select = c(Number.Policy.Passed, Percent.Policy.Passed.Of.All.Proposed, Number.of.Allergen.Species, Smoking.prevalence..as.a.percentage.of.the.population., E_TOTPOP, High.Pollen.State., Percent.Properties.Medium.Or.High.Fire.Risk, EP_UNINSUR, Prevalence.Lung.Cancer))

tree <- rpart(RT_Limited_Variables_Train$Prevalence.Lung.Cancer ~ ., data=RT_Limited_Variables_Train, control=rpart.control(cp=.0001))
printcp(tree)
best <- tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"]

#produce a pruned tree based on the best cp value
pruned_tree <- prune(tree, cp=best)


#plot the pruned tree
prp(pruned_tree,
    faclen = 0,    # Use full names for factor labels
    extra = 1,     # Display the number of observations for each terminal node
    roundint = F,  # Don't round to integers in output
    digits = 1,    # Display 1 decimal place in output
    varlen = 20,   # Increase variable name length for readability
    branch.lty = 2, # Dashed branch lines for better visibility
    tweak = .4,   # Adjust this value to increase/decrease the font size
    box.palette = "Blues", # Change the color palette (use a different name or color)
    yesno = 1,     # Show "yes" or "no" at nodes
    type = 2,      # Use "compact" format
    cex = 0.8,     # Adjust this value to increase/decrease the font size
    mar = c(0, 0, 2, 0), # Adjust the plot margins as needed
    main = "Pruned Regression Tree") # Set a main title for the plot

# Calculate the R-squared value

# Make predictions (change "new_data" to your new data for prediction)
# new_data <- data.frame(...)  # Replace with your new data
predictions <- predict(pruned_tree, newdata = LC_Test)

Test_observations <- LC_Test$Prevalence.Lung.Cancer
# Evaluate the model's performance (e.g., using confusion matrix, ROC, etc.)
# Calculate the R-squared value

rsquared <- R2(pred = predictions, obs = Test_observations)
print(paste("R-squared:", rsquared))

# RMSE
RMSE(predictions,Test_observations)


```

http://www.sthda.com/english/wiki/wiki.php?id_contents=7940
https://rpubs.com/corey_sparks/539484

```{r}
############Clustering###########
#Data does need to be scaled
DF_No_Predictors <- subset(DF, select = -c(Prevalence.Pediatric.Asthma, Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))

Predictors <- subset(DF, select = c(Prevalence.Pediatric.Asthma, Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))

DF_Clustering <- scale(DF_No_Predictors, center=TRUE, scale=TRUE)
colnames(DF_Clustering) <- colnames(DF_No_Predictors)

DF_Clustering_Final <- cbind(DF_Clustering, Predictors)

set.seed(123)
#training test
n <- nrow(DF)
sample_indices <- sample(1:n, size = n, replace = FALSE)
train_size <- round(0.7 * n)
test_size <- n - train_size
train_data_CL <- DF_Clustering_Final[sample_indices[1:train_size], ]
test_data_CL <- DF_Clustering_Final[sample_indices[(train_size + 1):n], ]

train <-subset(train_data_CL, select = -c(Prevalence.Pediatric.Asthma, Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
train_response <- subset(train_data_CL, select = c(Prevalence.Pediatric.Asthma, Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
test <- subset(test_data_CL, select = -c(Prevalence.Pediatric.Asthma, Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
test_response <- subset(test_data_CL, select = c(Prevalence.Pediatric.Asthma, Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))

# Set the number of clusters (e.g., k = 3)
k <- 5

# Run k-means clustering
kmeans_result <- kmeans(train, centers = k, nstart = 25)  # nstart controls the number of initial random configurations

# Get cluster assignments for each data point
cluster_assignments <- kmeans_result$cluster

# View cluster assignments
print(cluster_assignments)

cluster_df <- data.frame(Cluster = cluster_assignments, Response = train_response)
library(ggplot2)

ggplot(cluster_df, aes(x = Cluster, y = train_response$Prevalence.COPD, color = as.factor(Cluster))) +
  geom_jitter(width = 0.2) +  # Add jitter to points for better visibility
  labs(x = "Cluster", y = "Response Variable") +
  ggtitle("Cluster Assignments vs. Response Variable")


test_clusters <- kmeans(test, centers = kmeans_result$centers, nstart = 25)
test_cluster_assignments <- test_clusters$cluster
cluster_df2 <- data.frame(Cluster = test_cluster_assignments, Response = test_response)

ggplot(cluster_df2, aes(x = test_cluster_assignments, y = test_response$Prevalence.COPD, color = as.factor(Cluster))) +
  geom_jitter(width = 0.2) +  # Add jitter to points for better visibility
  labs(x = "Cluster", y = "Response Variable") +
  ggtitle("Cluster Assignments vs. Response Variable")



#### DBSCAN

# Install and load the dbscan package
library(dbscan)

# Assuming your dataset is named "data" and contains the features for clustering
# Specify your eps and minPts values
eps <- 3
minPts <- 2

# Perform DBSCAN clustering
dbscan_result <- dbscan(train, eps = eps, MinPts = minPts, )

# Access the cluster assignments
cluster_assignments <- dbscan_result$cluster

cluster_df <- data.frame(Cluster = cluster_assignments, Response = train_response)

ggplot(cluster_df, aes(x = Cluster, y = train_response$Prevalence.Lung.Cancer, color = as.factor(Cluster))) +
  geom_jitter(width = 0.2) +  # Add jitter to points for better visibility
  labs(x = "Cluster", y = "Response Variable") +
  ggtitle("Cluster Assignments vs. Response Variable")


```


```{r}
###########Gradient Boosting###########


# Define the XGBoost parameters
params = {
    'objective': 'reg:squarederror',  # For regression tasks
    'eval_metric': 'rmse',            # Root Mean Squared Error as the evaluation metric
    'max_depth': 6,                   # Maximum depth of the tree
    'learning_rate': 0.1,             # Learning rate
    'n_estimators': 100,              # Number of boosting rounds (trees)
    'subsample': 0.8,                # Fraction of samples used for tree building
    'colsample_bytree': 0.8,         # Fraction of features used for tree building
    'random_state': 42               # Random seed for reproducibility
}

```


```{r}
###########Gradient Boosting###########

set.seed(123)
#training test
n <- nrow(DF)
sample_indices <- sample(1:n, size = n, replace = FALSE)
train_size <- round(0.7 * n)
test_size <- n - train_size
train_data_NL <- DF[sample_indices[1:train_size], ]
test_data_NL <- DF[sample_indices[(train_size + 1):n], ]

#get data split apart
COPD_Train <-subset(train_data_NL, select = -c(Prevalence.Pediatric.Asthma,Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
COPD_Train_response <-subset(train_data_NL, select = c(Prevalence.COPD))

COPD_Test <- subset(test_data_NL, select = -c(Prevalence.Pediatric.Asthma,Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
COPD_Test_response <- subset(test_data_NL, select = c(Prevalence.COPD))


PA_Train <- subset(train_data_NL, select = -c(Prevalence.Pediatric.Asthma,Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
PA_Train_response <-subset(train_data_NL, select = c(Prevalence.Pediatric.Asthma))
PA_Test <- subset(test_data_NL, select = -c(Prevalence.Pediatric.Asthma,Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
PA_Test_response <-subset(test_data_NL, select = c(Prevalence.Pediatric.Asthma))


AA_Train <- subset(train_data_NL, select = -c(Prevalence.Pediatric.Asthma,Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
AA_Train_response<-subset(train_data_NL, select = c(Prevalence.Adult.Asthma))
AA_Test <- subset(test_data_NL, select = -c(Prevalence.Pediatric.Asthma,Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
AA_Test_response<-subset(test_data_NL, select = c(Prevalence.Adult.Asthma))

CLD_Train <- subset(train_data_NL, select = -c(Prevalence.Pediatric.Asthma,Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
CLD_Train_response<-subset(train_data_NL, select = c(Prevalence.Adult.Chronic.Lung.Disease))
CLD_Test <- subset(test_data_NL, select = -c(Prevalence.Pediatric.Asthma,Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
CLD_Test_response<-subset(test_data_NL, select = c(Prevalence.Adult.Chronic.Lung.Disease))

LC_Train <- subset(train_data_NL, select = -c(Prevalence.Pediatric.Asthma,Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
LC_Train_response<-subset(train_data_NL, select = c(Prevalence.Lung.Cancer))

LC_Test <- subset(test_data_NL, select = -c(Prevalence.Pediatric.Asthma,Prevalence.COPD, Prevalence.Adult.Chronic.Lung.Disease, Prevalence.Lung.Cancer, Prevalence.Adult.Asthma))
LC_Test_response<-subset(test_data_NL, select = c(Prevalence.Lung.Cancer))
```


```{r}
# Define the XGBoost parameters
params <- list(
  objective = 'reg:squarederror',
  eval_metric = 'rmse',
  max_depth = 6,
  learning_rate = 0.1,
  subsample = 0.8,
  colsample_bytree = 0.8,
  random_state = 42
)
```


```{r}
X_train <- as.matrix(COPD_Train)
y_train <- as.matrix(COPD_Train_response$Prevalence.COPD)
xgb_model <- xgboost(data = X_train, label = y_train, params = params, nrounds = 200)


X_test <- as.matrix(COPD_Test)
y_test <- as.matrix(COPD_Test_response$Prevalence.COPD)

# Make predictions on the test data
predictions <- predict(xgb_model, newdata=X_test)

# Calculate RMSE and R-squared
rmse <- RMSE(predictions,y_test)
r2 = R2(y_test, predictions)

print(rmse)
print(r2)
```


```{r}
#######Gain represents fractional contribution of each feature to the model based on the total gain of this feature's splits. Higher percentage means a more important predictive feature.

importance_scores <- xgb.importance(model = xgb_model)
importance_df <- as.data.frame(importance_scores)
excel_file_path <- "Gradient_Boost_Feature_Importance_COPD.xlsx"
write_xlsx(importance_df, path = excel_file_path)
```


```{r}




```


